---
title: "Bank Marketing"
author: "Vanessa okosun"
date: '2023-10-11'
output: html_document
---


```{r}
knitr::opts_chunk$set(echo = TRUE)
install.packages("GGally")
install.packages("scales")
install.packages("ggcorrplot")
install.packages("fastDummies")
package.installed <- c("readr", "tidyverse", "ggplot2")
install.packages(package.installed, dependencies = TRUE)

library(GGally)
```

```{r}
library(readr)
library(tidyverse)
```


```{r}
library(ggplot2)
library(caret)
```


```{r}
library(dplyr)
library(gridExtra)
```

```{r}
library(corrplot) #correlation plot
```

```{r}
library(ggcorrplot) #correlation for categorical
library(fastDummies) # create dummies variable
```


BUSINESS QUESTION
Bank wants to start a new marketing campaign but it wants to determine which prospect to target for its banking service

The objective is to use data mining techniques to classify which prospects are likely to apply for loan

LOAD DATA
```{r}
BMarketing <- read.csv('/Users/vanessaokosun/Documents/Job Stuff/Portfolio Projects/Bank Marketing/Banking Marketing Data/data.csv')
```

DATA STRUCTURE
Banking Marketing dataset has 41188 observations and 21 variables
```{r}
names(BMarketing)
```

```{r}
str(BMarketing)
```

```{r}
#dim(BMarketing)
BMarketing.na <- na.omit(BMarketing)
dim(BMarketing.na)
```


Change response var y to num
```{r}
BMarketing$y = ifelse(BMarketing$y=='yes',1,0)
head(BMarketing)
```

Basic Descriptive stats
Stats for numeric variables
```{r}
library(psych)
```

```{r}
summary(BMarketing)
```

```{r}
describe(BMarketing)
```
 
```{r}
#install.packages("skimr")
library(skimr)

skim(BMarketing)
```

Stats for categorical variables
```{r}
install.packages("jmv")
library("jmv")
```

```{r}
BMarketing %>%
  select(job, marital, education, default, housing, loan, contact, month, day_of_week, poutcome) %>%
  descriptives(freq = TRUE)
```

VISUALIZATION
```{r}
#
ggplot(BMarketing, aes(x = marital, fill = default))+ 
  geom_bar(stat="count", width=0.7)+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Default by marital status")
```

```{r}
ggplot(BMarketing, aes(x = education, fill = poutcome))+ 
  geom_bar(stat="count", width=0.7)+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Previous Campaign Outcome by education")
```

```{r}
ggplot(BMarketing, aes(x = job, fill = job))+ 
  geom_bar(stat="count", width=0.7)+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Previous Campaign Outcome by job")
```

CORRELATION ANALYSIS
CREATE NUMERIC DATASET FOR CORRELATION
```{r}
BMarketing.numer <- Filter(is.numeric, BMarketing)

names(BMarketing.numer)
```

```{r}
cor(BMarketing.numer, use = "complete.obs", method = "pearson")
```

```{r}
library(corrplot)
corrplot(cor(BMarketing.numer), method = "color")
```

```{r}
library(ggcorrplot)
ggcorrplot(cor(BMarketing.numer),
            hc.order = TRUE, 
           type = "lower",
           lab = TRUE)
```

Variables to exclude
```{r}
to.drop <- c("pdays", "nr.employed", "age", "cons.conf.idx", "campaign", "euribor3m")
to.drop
```

Dataset BMarketing.V1 exclude correlated variables
```{r}
to.drop <- c("pdays", "nr.employed", "age", "cons.conf.idx", "campaign", "euribor3m")

BMarketing.V1 <- BMarketing[ , -which(names(BMarketing) %in% to.drop)]

names(BMarketing.V1)
```

```{r}
write.csv(BMarketing,'BMarketing.V1.csv')


str(BMarketing.V1)
```

Transform character variables to factors
BMarketing.V2 dataset contains categorical variable that have been factorized
```{r}
#convert to factor
library(data.table)
```

```{r}
copy.BMarketing.V1 <- copy(BMarketing.V1)

cols <- c("job", "marital", "education", "default", "housing", "loan","contact", "month", "day_of_week", "poutcome", "y" )
BMarketing.V2 <- BMarketing.V1[cols] <- lapply(BMarketing.V1[cols], factor)

#BMarketing.V2$y <- ifelse(BMarketing.V2$y== 1, "yes", "no")

#BMarketing.V2$y <- as.factor(BMarketing.V2$y)

write.csv(BMarketing.V1,'BMarketing.V2.csv')

str(BMarketing.V2)
```

Reduce dimensionlatity of job
```{r}
# recoding job
BMarketing.V2$job1 <- case_when(
      BMarketing.V2$job == "unknown" ~ "unconventional",
      BMarketing.V2$job == "unemployed" ~ "unconventional",
      BMarketing.V2$job == "entrepreneur" ~ "unconventional",
      BMarketing.V2$job == "retired" ~ "unconventional",
      BMarketing.V2$job == "housemaid" ~ "unconventional",
      BMarketing.V2$job == "self-employed" ~ "unconventional",
      BMarketing.V2$job == "student" ~ "unconventional",
      BMarketing.V2$job == "blue-collard" ~ "blue-collar",
      BMarketing.V2$job == "technician" ~ "blue-collar",
      BMarketing.V2$job == "services" ~ "blue-collar",
      TRUE ~ "white-collar")
                                
 BMarketing.V2$job1 <- as.factor(BMarketing.V2$job1)                           
 str(BMarketing.V2)
```

Recode marital
```{r}
# Recode marital status
BMarketing.V2$marital1 <- case_when(BMarketing.V2$marital == "divorced" ~ "single",
                            BMarketing.V2$marital == "unknown" ~ "single",
                            BMarketing.V2$marital == "single" ~ "single",
                            TRUE ~ "married")

BMarketing.V2$marital1 <- as.factor(BMarketing.V2$marital1)

BMarketing.V2 <- as.data.frame(BMarketing.V2)

str(BMarketing.V2)
```

recode default
```{r}
# Recode default
BMarketing.V2$default1 <- case_when(BMarketing.V2$default == "unknown" ~ "no",
                            BMarketing.V2$default == "no" ~ "no",
                            TRUE ~ "yes")

BMarketing.V2$default1 <- as.factor(BMarketing.V2$default1)
```

recode housing to merge unknown to no
```{r}
BMarketing.V2$housing1 <- case_when(BMarketing.V2$housing == "unknown" ~ "no",
                            BMarketing.V2$housing == "no" ~ "no",
                            TRUE ~ "yes")

BMarketing.V2$housing1 <- as.factor(BMarketing.V2$housing1)

str(BMarketing.V2)
```

```{r}
pt1 <-ggplot(data=BMarketing.V2, aes(x=marital, fill=marital))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Marital")

pt2 <- ggplot(data=BMarketing.V2, aes(x= marital1, fill= marital1))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Recoded Marital")
grid.arrange(pt1,pt2, ncol=2)
```

recode loan
```{r}
BMarketing.V2$loan1 <- case_when(BMarketing.V2$loan == "unknown" ~ "no",
                               BMarketing.V2$loan == "no" ~ "no",
                               TRUE ~ "yes")
                    
BMarketing.V2$loan1 <- as.factor(BMarketing.V2$loan1)
str(BMarketing.V2)
```

Recode housing
```{r}
#recode Housing
BMarketing.V2$housing1 <- case_when(BMarketing.V2$housing == "unknown" ~ "no",
                            BMarketing.V2$housing == "no" ~ "no",
                            TRUE ~ "yes")
BMarketing.V2$housing1 <- as.factor(BMarketing.V2$housing)

str(BMarketing.V2)
```

Recode education
```{r}
BMarketing.V2$education1 <- case_when(
      BMarketing.V2$education == "unknown" ~ "noCollege",
      BMarketing.V2$education == "basic.4y" ~ "noCollege",
      BMarketing.V2$education == "basic.6y" ~ "noCollege",
      BMarketing.V2$education == "basic.9y" ~ "noCollege",
      BMarketing.V2$education == "illiterate" ~ "noCollege",
      BMarketing.V2$education == "high.school" ~ "noCollege",
      BMarketing.V2$education == "professional.course" ~ "noCollege",
      TRUE ~ "College.dg")
BMarketing.V2$education1 <- as.factor(BMarketing.V2$education1)

str(BMarketing.V2)
```

```{r}
table(BMarketing.V2$month)
```

```{r}
pt1 <-ggplot(data=BMarketing.V2, aes(x=education, fill= education))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Education")

pt2 <- ggplot(data=BMarketing.V2, aes(x= education1, fill= education1))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Recoded Education")
grid.arrange(pt1,pt2, ncol=2)
```

```{r}
#library(forcats)
pt1 <- ggplot(data=BMarketing.V2, aes(x=job1, fill=poutcome))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Loan Success by profession")

pt2 <- #library(forcats)
ggplot(data=BMarketing.V2, aes(x=education1, fill=poutcome))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Loan Success by education")
grid.arrange(pt1,pt2, ncol=2)
```

recoding month
```{r}
BMarketing.V2$month1 <- case_when(
      
      BMarketing.V2$month == "apr" ~ "Q2",
      BMarketing.V2$month == "may" ~ "Q2",
      BMarketing.V2$month == "jun" ~ "Q2",
      BMarketing.V2$month == "jul" ~ "Q3",
      BMarketing.V2$month == "aug" ~ "Q3",
      BMarketing.V2$month == "sep" ~ "Q3",
      TRUE ~ "Q4")
BMarketing.V2$month1 <- as.factor(BMarketing.V2$month1)

str(BMarketing.V2)
```

```{r}
# pt1 <-ggplot(data=BMarketing.V2, aes(x=month1, fill= y))+
#            geom_bar()+
#     theme(axis.text.x=element_text(angle=35,hjust=1))+
#     ggtitle("Current Campaign Result")
# 
# pt2 <- ggplot(data=BMarketing1, aes(x= poutcome, fill= poutcome))+
#            geom_bar()+
#     theme(axis.text.x=element_text(angle=35,hjust=1))+
#     ggtitle("Past Campaign Result")
# grid.arrange(pt1,pt2, ncol=2)
```

BMarketing.V3 remove original var that have been recoded
```{r}
# BMarketing.V3 <- within(BMarketing.V2, rm(job, marital,loan,housing,education,month, default))
# 
# write.csv(BMarketing.V2, "BMarketing.V3")
# 
# BMarketing.V3 <- as.data.frame(BMarketing.V3)
# str(BMarketing.V3)
```

Change order of col to put target var: y first
```{r}
# BMarketing.V3.1 <- BMarketing.V3[ ,c(4,1,2,3,5,6,7,8,9,10,11)]
# 
# write.csv(BMarketing.V3, "Bmarketing.V3.1")
# 
# BMarketing.V3.1$y <- ifelse(BMarketing.V3.1$y== 1, "yes", "no")
# BMarketing.V3.1$y  <- as.factor(BMarketing.V3.1$y)
# 
# 
# str(BMarketing.V3.1)
```

Dummy variables
```{r}
# library(fastDummies)
# 
# BMarketing.V3.1.dum <- dummy_cols(BMarketing.V3.1, select_columns = c("default1", "contact","day_of_week", "poutcome", "job1", "marital1", "loan1", "housing1", "education1", "month1"), remove_first_dummy = T)
# 
# BMarketing.V3.1.dum$y <- ifelse(BMarketing.V3.1.dum$y == "yes", 1, 0)
# 
# str(BMarketing.V3.1.dum)
# 
# BMarketing.V3.1.dum1 <- within(BMarketing.V3.1.dum, rm(default1, contact,day_of_week, poutcome,job1,marital1,loan1, housing1,education1, month1))
# 
# BMarketing.V3.1.dum1$y <- ifelse(BMarketing.V3.1.dum1$y== "yes", 1, 0)
# 
# #write.csv(BMarketing.V3.1.dum, "BMarketing.V3.dum1")
# 
# 
# 
# str(BMarketing.V3.1.dum1)
```

```{r}
# BMarketing.V3.1$y <- ifelse(BMarketing.V3.1$y== 1, "yes", "no")
# BMarketing.V3.1$y  <- as.factor(BMarketing.V3.1$y)
# 
# 
# str(BMarketing.V3.1)
```

```{r}
#prop.table(table(BMarketing1$y))

```

Proportion of approval:
87% rejected to 13% acceptance. Imbalance in the dataset, we need oversampling to balance the dataset to avoid bias in our model

MODELS

MODEL 1: KNN
```{r}
## BMarketing3 subset of numeric
# 
# str(BMarketing.numer)
# num.excl <- c("age", "campaign", "pdays", "emp.var.rate", "cons.price.idx", "euribor3m", "nr.employed")
# 
# BMarketing.numer1 <- BMarketing.numer[ , -which(names(BMarketing.numer) %in% num.excl)] #excl high corr var
# 
# str(BMarketing.numer1)
# 
# head(BMarketing.numer1)
```

```{r}

# library(class)
# library(caret)
# library(FNN)
# 
# set.seed(31)
# 
# #Partition the data
# BMk.train.index <- sample(row.names(BMarketing.V3.1.dum1), 0.6* dim(BMarketing.V3.1.dum1)[1])
# BMk.valid.index <- setdiff(row.names(BMarketing.V3.1.dum1), BMk.train.index)
# BMk.train <- BMarketing.V3.1.dum1[BMk.train.index, ]
# BMk.valid <- BMarketing.V3.1.dum1[BMk.valid.index, ]
# 
# class(BMk.valid$y)
# 
# BMk.train[, -1] <-predict( BMk.train[ ,-1])
# BMk.valid[, -1] <- predict( BMk.valid[, -1])
# BMarketing.V3.1.dum1[ , -1] <- predict( BMarketing.numer1[ , -1])
# options(scipen=999, digits=5)
# 
# str(BMk.train)

#initialized the normalized training, validation and full dataset
# BMk.train.norm <- BMk.train
# BMk.valid.norm <- BMk.valid
# BMarketing.numer1.norm <- BMarketing.numer1

# tail(BMarketing.numer1)

#scale the training, validation and full dataset
#norm.k.values <- preProcess(BMk.train[, 1:3], method = c("center","scale"))

#BMk.train.norm[, -4] <-predict(norm.k.values, BMk.train[ ,-4])
#BMk.valid.norm[, -4] <- predict(norm.k.values, BMk.valid[, -4])
#BMarketing.numer1.norm[ , -4] <- predict(norm.k.values, BMarketing.numer1[ , -4])
#options(scipen=999, digits=5)
                                          
#initialize k and accuracy
#k.accuracy.df <- data.frame(k= seq(1,10,1), RMSE=rep(0,10))

#knn for different k on validation

# #for(i in 1:10) {
#  # BM.knn.pred <- knn(BMk.train.norm[, 1:3],
#                      BMk.valid.norm[, 1:3], 
#                      cl = BMk.train.norm[, 4], k = i)
#  # k.accuracy.df[i, 2] <- confusionMatrix(factor(BM.knn.pred), factor(BMk.valid.norm[, 4]))$overall[1]
#}

#str(BMk.train.norm)
#str(BMk.train.norm)

#confusionMatrix(factor(BM.knn.pred), factor(BMk.valid.norm[, 4]))

# for(i in 1:10) {
#     BM.knn.pred <- class::knn(BMk.train.norm[ ,-4],
#                            test = BMk.valid.norm[, -4],
#                            cl = BMk.train.norm$y, k=i)
#     k.accuracy.df[i,2] <- RMSE(as.numeric(as.character(BM.knn.pred)), BMk.valid.norm$y)
# }
# 
# k.accuracy.df
# 
# newHouse.df <- data.frame(duration = 239, previous=1, cons.conf.idex=-50.8)
#newHouse1.df <- data.frame(duration = 189, previous=0, cons.conf.idx= -50.8)
```

For k= 5 accuracy is accuracy of 0.9027. We get better accuracy with k=5, with k been in the middle to avoid overfitting
```{r}
# confusionMatrix(factor(BM.knn.pred), factor(BMk.valid.norm[, 4]))
```

MODEL 2: LOGISTIC REGRESSION
```{r}
# str(BMarketing.V3.1)
```

partition data for logistic reg
```{r}
# set.seed(51)
# BMl.train.index <- sample(row.names(BMarketing.V3.1), 0.6* dim(BMarketing.V3.1)[1])
# BMl.valid.index <- setdiff(row.names(BMarketing.V3.1), BMl.train.index)
# BMl.train <- BMarketing.V3.1[BMl.train.index, ]
# BMl.valid <- BMarketing.V3.1[BMl.valid.index, ]
```

Logistic reg
```{r}
# BM.V3.1.logitlm <- glm(y ~ ., data=BMl.train, family=binomial(link="logit"), na.action = na.exclude)
# options(scipen=999)
# 
# summary(BM.V3.1.logitlm)
```

```{r}
# round(data.frame(summary(BM.V3.1.logitlm)$coefficients, odds = exp(coef(BM.V3.1.logitlm))),5)
```

with a cut off of 0.5, our model accurately predicted that our 5 record as no good prospect for term deposit
```{r}
# BM.V3.logit.reg.pred <- predict(BM.V3.1.logitlm, BMl.valid [, -1], type = "response")
# 
# # first 5 actual and predicted records
# data.frame(actual = BMl.valid$y[1:5], predicted = BM.V3.logit.reg.pred[1:5])
# 
```

```{r}
# library(gains)
# dim(BMl.valid)
# 
# length(BM.V3.logit.reg.pred)
# 
# BMl.valid$y <- ifelse(BMl.valid$y== "yes", 1, 0)
# 
# BM.gain <- gains(BMl.valid$y, BM.V3.logit.reg.pred, groups=10)
# # plot lift chart
# plot(c(0,BM.gain$cume.pct.of.total*sum(BMl.valid$y))~c(0,BM.gain$cume.obs),
# xlab="# cases", ylab="Cumulative", main="", type="l")
# lines(c(0,sum(BMl.valid$y))~c(0, dim(BMl.valid)[1]), lty=2)
```

```{r}
# # compute deciles and plot decile-wise chart
# heights <- BM.gain$mean.resp/mean( BMl.valid$y)
# midpoints <- barplot(heights, names.arg = BM.gain$depth, ylim = c(0, 9),
#                      xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# # add labels to columns
# text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```

Confusion matrix
```{r}
# library(caret)
# BM.V3.logit.reg.pred <- predict(BM.V3.1.logitlm, BMl.valid [, -1], type = "response")
# 
# BM.V3.logit.reg.pred2<-ifelse(BM.V3.logit.reg.pred > 0.5, 1, 0)
# 
# confusionMatrix (as.factor(BM.V3.logit.reg.pred2), as.factor(BMl.valid$y), positive = '1')
```

Naive Bayes
```{r}
# str(BMarketing.V3.1)
```

training and validation set for NB
```{r}
# set.seed(154)
# BMnb.train.index <- sample(row.names(BMarketing.V3.1), 0.6* dim(BMarketing.V3.1)[1])
# BMnb.valid.index <- setdiff(row.names(BMarketing.V3.1), BMnb.train.index)
# BMnb.train <- BMarketing.V3.1[BMnb.train.index, ]
# BMnb.valid <- BMarketing.V3.1[BMnb.valid.index, ]
```

fit the model
```{r}
# 
# library(e1071)
# BM.nb <- naiveBayes(y~ .,data=BMnb.train)
# BM.nb
```

```{r}
# pred.tr.class <- predict(BM.nb, newdata=BMnb.train, type="class")
# 
# confusionMatrix(pred.tr.class, BMnb.train$y)
# 
```

```{r}
# pred.vl.class <- predict(BM.nb, newdata=BMnb.valid, type="class")
# 
# pred.vl.prob <- predict(BM.nb, newdata=BMnb.valid, type="raw")
# 
# confusionMatrix(pred.vl.class, BMnb.valid$y)
```

```{r}

# library(gains)
# 
# str(BMarketing.V3.1)
# 
# BMarketing.V3.1$y <- ifelse(BMarketing.V3.1$y == 1, "yes","no")
# 
# BMarketing.V3.1$y <- as.factor(BMarketing.V3.1$y)
# 
# nb.gain <- gains(ifelse(BMnb.valid$y=="no",0,1), pred.vl.prob[,1], groups=100)
# 
# #nb.gain <- gains(pred.vl.prob[,1], groups=100)
# 
# plot(c(0,nb.gain$cume.pct.of.total*sum(BMnb.valid$y=="no"))~c(0,nb.gain$cume.obs), 
#      xlab="# cases", ylab="Cumulative", main="", type="l")
# lines(c(0,sum(BMnb.valid$y=="no"))~c(0, dim(BMnb.valid)[1]), lty=2)
```

```{r}
# compute deciles and plot decile-wise chart
# heights <- nb.gain$mean.resp/mean( BMnb.valid$y)
# midpoints <- barplot(heights, names.arg = nb.gain$depth, ylim = c(0, 9),
#                      xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# # add labels to columns
# text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```

OVERSAMPLING
```{r}
# str(BMarketing.V1)
```

```{r}
# str(bmarketing.v1)
```

Recode variable
```{r}
#recode marital

bmarketing.v1$marital <- case_when(bmarketing.v1$marital == "divorced" ~ "divorced",
                            bmarketing.v1$marital == "unknown" ~ "single",
                            bmarketing.v1$marital == "single" ~ "single",
                            TRUE ~ "married")

bmarketing.v1$marital <- as.factor(bmarketing.v1$marital)

bmarketing.v1 <- as.data.frame(bmarketing.v1)

str(bmarketing.v1)
```

```{r}
## recode loan
bmarketing.v1$loan <- case_when(bmarketing.v1$loan == "unknown" ~ "no",
                               bmarketing.v1$loan == "no" ~ "no",
                               TRUE ~ "yes")
                    
bmarketing.v1$loan <- as.factor(bmarketing.v1$loan)
str(bmarketing.v1)
```

```{r}
## recode default
bmarketing.v1$default <- case_when(bmarketing.v1$default == "unknown" ~ "no",
                            bmarketing.v1$default == "no" ~ "no",
                            TRUE ~ "yes")

bmarketing.v1$default <- as.factor(bmarketing.v1$default)


###recode job
bmarketing.v1$job <- case_when(
      bmarketing.v1$job == "unknown" ~ "unemployed",
      bmarketing.v1$job == "unemployed" ~ "unemployed",
      bmarketing.v1$job == "entrepreneur" ~ "entrepreneur",
      bmarketing.v1$job == "retired" ~ "retired",
      bmarketing.v1$job == "housemaid" ~ "housemaid",
      bmarketing.v1$job == "self-employed" ~ "self-employed",
      bmarketing.v1$job == "student" ~ "student",
      bmarketing.v1$job == "blue-collard" ~ "blue-collar",
      bmarketing.v1$job == "technician" ~ "blue-collar",
      bmarketing.v1$job == "services" ~ "blue-collar",
      TRUE ~ "white-collar")
                                
 bmarketing.v1$job <- as.factor(bmarketing.v1$job) 
 
 str(bmarketing.v1)
```

```{r}
 #recode housing
 bmarketing.v1$housing <- case_when(bmarketing.v1$housing == "unknown" ~ "no",
                            bmarketing.v1$housing == "no" ~ "no",
                            TRUE ~ "yes")
bmarketing.v1$housing <- as.factor(bmarketing.v1$housing)

str(bmarketing.v1)
```

```{r}
## recoding education
bmarketing.v1$education <- case_when(
      bmarketing.v1$education == "unknown" ~ "noCollege",
      bmarketing.v1$education == "basic.4y" ~ "noCollege",
      bmarketing.v1$education == "basic.6y" ~ "noCollege",
      bmarketing.v1$education == "basic.9y" ~ "noCollege",
      bmarketing.v1$education == "illiterate" ~ "noCollege",
      bmarketing.v1$education == "high.school" ~ "noCollege",
      bmarketing.v1$education == "professional.course" ~ "noCollege",
      TRUE ~ "College.dg")
bmarketing.v1$education <- as.factor(bmarketing.v1$education)

str(bmarketing.v1)
```

```{r}
ggplot(data=bmarketing.v1, aes(x=education, fill=education))+
           geom_bar()+
    theme(axis.text.x=element_text(angle=35,hjust=1))+
    ggtitle("Recoded Education")
```

Proporting of reponse variable
```{r}
prop.table(table(bmarketing.v1$y))
```

```{r}
set.seed(456)

ind <- sample(2, nrow(bmarketing.v1), replace=TRUE, prob = c(0.6,0.4))

ov.train <- bmarketing.v1[ind==1, ]

ov.valid <- bmarketing.v1[ind==2,]
```

reponse prop in train
```{r}
prop.table(table(ov.train$y))
```

model random forest
```{r}
#install.packages("randomForest")
library(caret)
library(e1071)
library(ggplot2)
library(randomForest)
```

```{r}
rftrain <- randomForest(y~ ., data= ov.train)
confusionMatrix(predict(rftrain, ov.valid), ov.valid$y, positive='1')
```

Oversampling to test random forest model again
```{r}
#install.packages("ROSE")
library(ROSE)
```

```{r}
set.seed(455)

ind <- sample(2, nrow(bmarketing.v1), replace=TRUE, prob = c(0.6,0.4))

ov.train1 <- bmarketing.v1[ind==1, ]

ov.valid1 <- bmarketing.v1[ind==2,]


over <- ovun.sample(y~., data=ov.train1, method = "over", N=43652)$data

## prediction with random forest using non balance data


rftrain1 <- randomForest(y~., data=ov.train1)

confusionMatrix(predict(rftrain1, ov.valid1), ov.valid1$y, positive='1')
```

```{r}
## confusion matrix with balance dataset

rfover <- randomForest(y~., data=over)

confusionMatrix(predict(rfover, ov.valid1), ov.valid1$y, positive='1')
```

```{r}
str(ov.train1)
```

```{r}
str(ov.valid1)
```

```{r}
table(ov.train1$y)
```

```{r}
table(over$y)
```

```{r}
str(over)
```

Logistic regression using balance dataset
```{r}
train1.glm <- glm(y~., data=ov.train1,family=binomial(link="logit"), na.action = na.exclude)
summary(train1.glm)
```

```{r}
train1.glm.pred <- predict(train1.glm, ov.valid1[, -15], type= "response")

train1.glm.pred2<-ifelse(train1.glm.pred > 0.5, 1, 0)

confusionMatrix(as.factor(train1.glm.pred2), as.factor(ov.valid1$y), positive = '1')
```

```{r}
# summary(lregtrain1)
```

Logistic regression using balance dataset
```{r}
# train1.glm <- glm(y~., data=ov.train1,family=binomial(link="logit"), na.action = na.exclude)
# summary(train1.glm)
# 
# train1.glm.pred <- predict(train1.glm, ov.valid1[, -15], type= "response")
# 
# train1.glm.pred2<-ifelse(train1.glm.pred > 0.5, 1, 0)
# 
# confusionMatrix(as.factor(train1.glm.pred2), as.factor(ov.valid1$y), positive = '1')
# 



#############################################

# over.train1.glm <- glm(y~., data=over.1, family=binomial(link="logit"), na.action = na.exclude)
# 
# over.train1.glm.pred <- predict(over.train1.glm, ov.valid.1[, -14], type= "response")
# 
# over.train1.glm.pred2<-ifelse(over.train1.glm.pred > 0.5, 1, 0)
# 
# confusionMatrix(as.factor(over.train1.glm.pred2), as.factor(ov.valid.1$y), positive = '1')

# over$y <- ifelse(over$y == 1, 'yes', 'no')
# over$y <- as.factor(over$y )
# 
# str(over)
# ov.over.1 <- within(over, rm(default))
# ov.valid.1 <- within(ov.valid1, rm(default))
```

Naive Bayes
```{r}
set.seed(950)

ind1 <- sample(2, nrow(bmarketing.v1), replace=TRUE, prob = c(0.6,0.4))

nb.train1 <- bmarketing.v1[ind==1, ]

nb.valid1 <- bmarketing.v1[ind==2,]


nb.over <- ovun.sample(y~., data=nb.train1, method = "over", N=43652)$data


## Naive Bayes with imbalance data
train.nbayes <- naiveBayes(y~., data= nb.train1)

nb.pred <- predict(train.nbayes, newdata= nb.valid1, type= "class")

confusionMatrix(nb.pred, nb.valid1$y, positive ='1')
```

```{r}
dim(nb.train1)
```

```{r}
dim(nb.valid1)
```

```{r}
dim(nb.over)
```

```{r}
length(nb.pred)
```

```{r}
length(nb.valid1)
```

```{r}
# BM.nb <- naiveBayes(y~ .,data=BMnb.train)
# BM.nb
# pred.tr.class <- predict(BM.nb, newdata=BMnb.train, type="class")
# 
# confusionMatrix(pred.tr.class, BMnb.train$y)
```

```{r}
ov.train.nbayes <- naiveBayes(y~., data= nb.over)

ov.nb.pred <- predict(ov.train.nbayes, newdata= nb.valid1, type= "class")

confusionMatrix(ov.nb.pred, nb.valid1$y, positive ='1')
```

```{r}

```
