---
title: "Drugs Relationship on Depression"
author: "Vanessa Okosun"
date: "2023-02-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

LOADING Data
```{r}
#install.packages("GGally")
#install.packages("gridExtra")
#install.packages("corrplot")
#install.packages("ggcorrplot")
#install.packages("fastDummies")

library(readxl)
library(dplyr)
library(GGally)
library(readr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(corrplot) #correlation plot
library(ggcorrplot) #correlation for categorical
library(fastDummies) # create dummies variable

DrugOnDep <- read_excel("C:\\Users\\vanes\\iCloudDrive\\Documents\\Grad school\\BDA 650\\Dataset\\BDA_OKOSUN.V_Data.xlsx")

DrugOnDep[] <- lapply(DrugOnDep, as.numeric)
DrugOnDep <- data.frame(DrugOnDep)

View(DrugOnDep)
```

DATA STRUCTURE 
```{r}
str(DrugOnDep)
summary(DrugOnDep)

```

GRAPHICAL SUMMARIES PT 1 
```{r}
#Lots of children under one that are included in this dataset 
hist(DrugOnDep$AGEYR)

```

CLEANING DATA
```{r}
#remove unimportant columns
#DrugOnDep <- subset(DrugOnDep, select = -c(RIDEXAGM))

#reduce values in race origin level 0 - other races ;1 - black non-hispanic
DrugOnDep$Race <- ifelse(DrugOnDep$Race==4,1,0 )

#reduce values in Education level 0 - highschool equiv or less than; 1 - college and above
DrugOnDep$EDUC <- ifelse(DrugOnDep$EDUC==4 | DrugOnDep$EDUC==5, 1,0 )

#reduce values in Marital level 0 - not married; 1 - married
DrugOnDep$MARTL <- ifelse(DrugOnDep$MARTL==1, 1,0 )

#reduce values in Income level; 0- under $20,000, 1 - over $20,000
DrugOnDep$INCOME <- ifelse(DrugOnDep$INCOME<= 4 | DrugOnDep$INCOME == 13, 0, 1 )


```

```{r}
DrugOnDepFinal<- DrugOnDep
str(DrugOnDepFinal)
view(DrugOnDepFinal)
```



DATA STRUCTURE 
```{r}
#install.packages("skimr")
library(skimr)

skim(DrugOnDepFinal)
```

Correlation Tests
```{r}
library(ggcorrplot)
cor(DrugOnDepFinal, use = "complete.obs", method = "pearson")
  
library(corrplot)
corrplot(cor(DrugOnDepFinal), method = "color")
```




GRAPHICAL SUMMARIES PT 1 
```{r}
library(ggplot2)
#new age distribution
hist(DrugOnDepFinal$AGEYR)

#count of race (1 -	Mexican American, 2	- Other Hispanic, 3-Non-Hispanic White, 4-Non-Hispanic Black, 5-Other Race - Including Multi-Racial)
ggplot(DrugOnDepFinal,aes(x=factor(Race)))+ geom_bar(position="dodge")

#count of gender (1-MALE, 2-Female)
ggplot(DrugOnDepFinal,aes(x=factor(GENDER)))+ geom_bar(position="dodge")

#count of education
ggplot(DrugOnDepFinal,aes(x=factor(EDUC)))+ geom_bar(position="dodge")

#count of Marital status
ggplot(DrugOnDepFinal,aes(x=factor(MARTL)))+ geom_bar(position="dodge")

```

GRAPHICAL SUMMARIES PT 2
```{r}
#Number of people depressed 
ggplot(DrugOnDepFinal,aes(x=factor(Depressed)))+ geom_bar(position="dodge")

#Number of people with drug use
ggplot(DrugOnDepFinal,aes(x=factor(Drug_user)))+ geom_bar(position="dodge")


```


LOGISTIC REGRESSION - MODEL 1 
```{r}
library(AICcmodavg)

#partition data 
set.seed(2)
train.index <- sample(c(1:dim(DrugOnDepFinal)[1]), dim(DrugOnDepFinal)[1]*0.7)
train.df <- DrugOnDepFinal[train.index, ]
valid.df <- DrugOnDepFinal[-train.index, ]

#run logistic regression
logit.reg <- glm(Drug_user~Depressed+ GENDER+ Race+ EDUC+ MARTL+ POVRAT, data = train.df, family = 'binomial')
options(scipen=999)
summary(logit.reg)

exp(coef(logit.reg))


```

MODEL 2
```{r}
logit.reg2 <- glm(Drug_user~Depressed+ GENDER+ MARTL+ Race, data = train.df, family = 'binomial')
options(scipen=999)
summary(logit.reg2)

exp(coef(logit.reg2))

```
MODEL Comparison
K: The number of parameters in the model.
AICc: The AIC value of the model. model that explains the most variation in the data has lower value 
Delta_AICc: The difference between the AIC of the best model compared to the current model being compared.
AICcWt: The proportion of the total predictive power that can be found in the model.
Cum.Wt: The cumulative sum of the AIC weights.
LL: The log-likelihood of the model. This tells us how likely the model is, given the data we used.

MODEL 2 Offers a better fit 
```{r}
#List of models 
models <- list(logit.reg, logit.reg2)

#model names 
mod.names <- c('dep.gen.rac.educ.martl.pov','dep.gen.martl' )

#Calculate AIC 
aictab(cand.set = models, modnames = mod.names)
```

ACCESSING THE MODEL 
 McFadden’s R2, which ranges from 0 to just under 1. Values close to 0 indicate that the model has no predictive power. In practice, values over 0.40 indicate that a model fits the data very well.
```{r}
library(caret)

#PREDICTIVE POWER
pscl::pR2(logit.reg2)["McFadden"]
pscl::pR2(logit.reg)["McFadden"]

#COMPUTE IMPORTANCE OF EACH VARIABLE 
caret::varImp(logit.reg)

#VIF - VALUE OVER 5 IS SEVERE multicollinearity
car::vif(logit.reg2)
```

MAKE Predictions

a Depressed status of “Yes” has a probability of being a drug user of .3796. Conversely, an individual with the same gender, race, martl, educ, and poverty rate but with a depression status of “No” has a probability of being a drug user of 0.2414
```{r}

#define two individuals
new <- data.frame(GENDER = 2, Race = 1, MARTL = 0, EDUC = 0, POVRAT = 2.17, Depressed = c(1, 0))

#Predicted probability 
predict(logit.reg2, new, type = "response")

```

That metric ranges from 0.50 to 1.00, and values above 0.80 indicate that the model does a good job in discriminating between the two categories which comprise our target variable. 
```{r}

library(pROC)
# Compute AUC for predicting Class with the variable CreditHistory.Critical
f1 = roc(Drug_user ~ Depressed, data=valid.df) 
plot(f1, col="red")

library(ROCR)
# Compute AUC for predicting Class with the model
prob <- predict(logit.reg2, newdata=valid.df, type="response")
pred <- prediction(prob, valid.df$Drug_user)
perf <- performance(pred, measure = "tpr", x.measure = "fpr")
plot(perf)
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]
auc


```

MODEL DIAGNOSTIC
```{r}
Predicted <- predict(logit.reg2, valid.df, type = 'response')

u <- union(Predicted, valid.df$Drug_user)
t <- table(factor(Predicted, u),factor(valid.df$Drug_user, u))

confusionMatrix(t)

class
```


LIFT CHART & DECILE-WISE LIFT CHART
```{r}
#install.packages('gains')
library(caret)
library(gains)
gain <- gains(valid.df$Drug_user, Predicted, groups= 100)

#LIFT CHART 
plot(c(0, gain$cume.pct.of.total*sum(valid.df$Drug_user))~c(0,gain$cume.obs),
xlab="# cases", ylab="Cumulative", main="", type="l")
lines(c(0,sum(valid.df$Drug_user))~c(0, dim(valid.df)[1]), lty=2)

```


#Simultaneous Equations Model#

the endogenous variables are  $Depression$ and  $Druguser$. 

the exogenous variables are the indicator variables for demographics of the user.

the identification variable for the drug user equation are gender, martial status and income , which will only show up in the depression equation but not in the drug user equation; 

the identification variables for the depression equation will be  drug prices, as none of them show up in the drug user equation.


```{r}
#Creating Drug Prices variables 
DrugOnDepFinal$Marij_price <- ifelse(DrugOnDepFinal$DUQ200=="1",7.38, 0)
DrugOnDepFinal$Cocaine_price <- ifelse(DrugOnDepFinal$DUQ250=="1",84.00, 0)
DrugOnDepFinal$Heroin_price <- ifelse(DrugOnDepFinal$DUQ290=="1",69.00, 0)
DrugOnDepFinal$Meth_price <- ifelse(DrugOnDepFinal$DUQ330=="1",40.00, 0)

view(DrugOnDepFinal)

```


```{r}
#install.packages('systemfit')
library(systemfit)

DrugOnDep.DU <- Drug_user~Depressed+Marij_price+Cocaine_price+Heroin_price+Meth_price
DrugOnDep.Dep <- Depressed ~ Drug_user+GENDER+MARTL+INCOME

DrugOnDep.eqs <- list(DrugOnDep.DU,DrugOnDep.Dep)
DrugOnDep.ivs <- ~Marij_price+Cocaine_price+Heroin_price+Meth_price+GENDER+MARTL+INCOME
DrugOnDep2SLS <-systemfit(DrugOnDep.eqs, method= "2SLS", 
                          inst=DrugOnDep.ivs, data= DrugOnDepFinal)

summary(DrugOnDep2SLS)

```


```{r}
DrugOnDep3SLS <- systemfit(DrugOnDep.eqs, method="3SLS", 
              inst=DrugOnDep.ivs, data=DrugOnDepFinal)
summary(DrugOnDep3SLS)
```

Hausman specification test
```{r}
h <- hausman.systemfit(DrugOnDep2SLS, DrugOnDep3SLS)
print(h)
```
X and Y (Depression and Drug user) are dependent of each other. Each variable can influence the other. This model helps correct for simultaneity
 
Based on the 2SLS model, the price of the 4 different drugs(marijuana, heroin, cocaine and meth) are very significant in the prediction of being a drug user.

In comparison, gender, marital status and income are also significant factor in the influence of depression.

Hausman specification test tells me to reject the 3SLS because the 2SLS is better. 

